### 2020-08-27 thesis/thesis.tex
change intro (andy's comments)

\SetAcknowledgeText{
For providing vital guidance and parts of the code, I thank Dr. Bronwyn Rayfield. For providing help with computing resources parts of the code, I thank Dr. Guillaume Larocque. For providing advice on STSIM implementation and code, I thank Dr. Colin Daniel and Dr. Leonardo Frid. I thank Dr. Guillaume Larocque and Dr. Yann Boulanger for their help in providing the LANDIS outputs and editorial help for the associated appendix. \\
\indent I thank all students, staff and organizations involved in the design and realisation of the workshop. I thank Aurélie Sierra for her invaluable guidance and counsel. I thank all students and staff facilitators: Madeleine Gauthier, Mikayla Salmon-Beitel and Violette Bertrand. I thank Geneviève Poirier and Mariève Lafontaine for their participation in the design, planning and organization of the workshop. I also thank Renée Gagnon for her precious help securing a location for the workshop.
\indent For providing essential funding for this project, I thank the Fonds de recherche du Québec – Nature et technologies (FRQNT). For providing logistical support I thank Dr. Helen Elina and Mr. Philippe Auzel. \\
\indent For providing key information on land use data, I thank Martin Joly, Marie-Josée Côté and Benoit Jobin. \\
\indent I thank my supervisor, Prof. Andrew Gonzalez, for his support and for believing in my abilities every step of the way. I also thank members of my committee, Prof. Brian Leung and Prof. Elena Bennett for their timely and wise council at every committee meeting. I thank all my fellow grad students and especially Alex Arkilanian for all the help revising proposals, discussion ideas, and providing key support. \\
\indent I thank my mother Laura, my father Philippe, and Talia for their unwavering support. Thank you to Blanka and Peter for sheltering me and putting up with my writing process during the thick of the pandemic. \\
}%
\Acknowledge%

@@

% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetIntroName{General Introduction}

\SetIntroText{

Space is a finite resource: there is a limit as to how much land is available for different human activities. How communities manage and govern space is a reflection of the choices and trade offs made by different people and organizations at different spatial and temporal scales, making the study of land systems a social, historical, and ecological discipline. The political process by which those trade-offs are resolved is of vital importance for biodiversity and ecosystem function. This process ultimately determines and regulates land use and land cover change (LUC): if and how the resources held on the land are exploited, transformed or conserved. LUC is an important cause of habitat loss, one of the most important threats to biodiversity and ecosystem function, reducing the diversity and abundance of species across the globe \citep{newbold_global_2015}. It is estimated that LUC is responsible for more than one-third of animal extinctions between 1600 and 1992 \citep{centre_global_1992}. Moreover, a report from the IPBES (Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services) shows that about about 9\% of all described terrestrial species might suffer from habitat loss to the point that it threatens their long-term persistence \citep{ipbes_summary_2019}. Therefore, habitat loss compromises many key ecosystem services, such as the focus of this thesis, ecological connectivity. \\

Ecological connectivity is the extent to which the landscape supports the movements of organisms \citep{gonzalez_spatial_2017}, and is paramount for the resilience of both populations and ecosystem services \citep{mitchell_Montérégie_2015} in heterogeneous and fragmented landscapes. Land use changes such as urban sprawl can cause deforestation, fragmenting habitats, and slowly eroding ecological connectivity. Many urban landscapes are experiencing uncontrolled urban sprawl and have suffered losses in connectivity and ecosystem services in consequence. Examples include cities like Barcelona \citep{marulli_gis_2005}, New York City \citep{mcphearson_urban_2014}, and also Montréal \citep{dupras_urban_2015}.  The drivers behind those land use changes are complex and understanding them is an obstacle to conservation planning  \citep{worboys_connectivity_2010}, including connectivity conservation planning. \\
Connectivity conservation planning (CCP), refers to the enterprise that engages multiple actors such as academics, NGOs, governmental bodies at different scales and in a common goal to conserve the ecological connectivity of the landscape. CCP methods usually entails modelling connectivity of the landscape of interest and using a prioritization method to determine conservation priorities, and modify land use practices to conserve or improve connectivity. This study constitutes a contribution to an ongoing effort of CCP in Montérégie, in southern Québec. In the following, we introduce the study region and describe previous efforts of CCP. We then point out ways in which CCP methods can be improved and how this thesis contributes to doing so. \\ \vspace{2em}

@@

Montérégie is one of the 17 administrative regions of the canadian province of Quebec. It is situated southeast of the city of Montréal, Québec, on the south bank of the Saint Lawrence, and is therefore located in the ecoregion of the Saint Lawrence Lowlands. Its superficy is of \SI{11852}{\km\squared} and is home to about $1.5$ million inhabitant. Agriculture is the predominant land use in the region, and about half of this superficy is covered by agricultural land \citep{isq_quebec_2020}. Statistique Quebec describes that between 1995 and 2008, 8\% of natural habitat and 4\% of agricultural lands has been artificialised. These changes are spread over the 15 MRC (“Municipalite Regionales de Comte”) making up Montérégie. MRCs are essentially sub-administrative regions at the level of which land use orientations are decided through two important documents: the PDZA (Plan de Développement de la Zone Agricole) and the SAD (Schéma d'Aménagement et de Développement). These MRCs are made up of 177 municipalities. In addition to MRCs, two First Nation reserves are present in Montérégie. \\

The eastern part of Montérégie is also under the jurisdiction of the Greater Montréal Area (GMA). The ecological connectivity of the GMA and its benefits as a provider of ecosystem services has recently been assessed in a report to the Québec ministry of the environment (Rayfield 2018, unpublished). This study focused on identifying regions of highest connectivity, and therefore of highest priority for the conservation of biodiversity and ecosystem services. Other work by Rayfield et al. (2019, unpublished) has extended the analysis of connectivity to the whole of the Saint Lawrence Lowlands. Beyond connectivity the project of Montérégie Connection, spearheaded by an interdisciplinary team of researchers from McGill University, studied the provision of ecosystem services in the region and produced important knowledge for conservation planning. \\

Deriving insight from these contributions, multiple stakeholders have become mobilized to set connectivity conservation goals for the region. A number of Non-Governmental NGOs have designed corridor networks over their jurisdiction, with varying degrees of inputs from previous scientific work. For instance, Nature Action Québec (NAQ) and the Centre de la Nature du Mont Saint Hilaire (CDN), but also the Centre d'interprétation du milieu écologique du Haut-Richelieu (CIME), and Appalachian Corridors (CA), which has been pioneering the approach in the region (although their jurisdiction is at the margin of Montérégie, in Estrie). These associations are now beginning to work together to design science-driven connectivity conservation plans for the whole region, and are in need of decision tools in that regard. They are concerned with the social acceptability of conservation, and with resilience of their plan to future land use and climate change. \\ \vspace{2em}

@@

The maps produced by previous analyses represent a snapshot of the current state of connectivity in the region. But researchers are developing methods for including future land use and climate change impacts in their analyses. For example, \cite{albert_applying_2017}, used a simple land use and climate change model  to identify a network of protected habitat robust to climate change for a subset of Montérégie. This is an important progress, because not taking into account risks of land use change could lead to ill-informed conservation plans with low efficacy. However, connectivity conservation models rarely include both land use and climate change in their calculations \citep{costanza_landscape_2019}, which represent a limitation to CCP methods. \\

Similarly, current CCP methods often do not confront the results of the prioritization with the priorities perceived by stakeholders. There is a need for planning tools to allow for this confrontation to happen. Those tools need to be able to guide stakeholders through an iterative process of conservation and land use scenario building. Methods are also available to include stakeholder’s input into CCP. Those “participatory research methods” are designed to conduct the research process “with those people whose life-world and meaningful actions are under study” \citep{bergold_participatory_2012}. Those methods often rely on collecting data through a community-driven process during workshops. \\

In this thesis, we show an attempt to remedy those two limitations. In the first chapter, we combined a connectivity model using circuit theory with a land use and climate change model that uses a combination of statistical modelling and MCMC-based simulations. In the second chapter, we ran a stakeholder workshop to derive conservation priorities and use those results as inputs to derive simple, community driven, alternative conservation scenarios. We compare and contrast the results of these scenarios, and discuss the importance of considering stakeholder inputs to produce a resilient network of protected areas. \\

Next, we introduce the three key conceptual themes of the thesis: connectivity modelling, land use change modelling, and participatory methods. \\ \vspace{2em}

Connectivity is a pillar of spatial ecology, the study of how ecological dynamics and interactions are structured in space. Connectivity is well known to influence various aspects of populations and communities, such as dispersal, demographic rescue, or the spread of diseases. In conservation, connectivity is an approach of interest to counteract the effects of  land use and climate change, and ensure the persistence of species in a landscape. \\

Connectivity as a concept has been used loosely in ecology and conservation. In their textbook “Spatial Ecology and Conservation Modeling”, \cite{fletcher_spatial_2018} identify up to 11 different such variations or nuances on the concept of connectivity. For the focus of this thesis, it is important to do two important distinctions with regard to the connectivity concept. First, a distinction is often made between structural and functional connectivity. Structural connectivity emphasizes landscape structure and configuration, while functional connectivity attempts to capture species’s diverse response to landscape’s configuration, and pay more attention to species-specific ways of moving across the landscape. Within functional connectivity another distinction is to be made between potential and realized connectivity \citep{calabrese_comparison_2004}. Potential connectivity estimates how much a landscape could facilitate or impedes animal movements, based on indirect information, such as species-specific models and expert knowledge. Realized connectivity, on the other hand, relies on observational data to measure animal movement directly to quantify connectivity. This thesis focuses on modelling potential functional modeling. \\

Connectivity modelling approaches depend on the level - population, community - and the scale - individual, patch or  landscape - that connectivity is considered, and therefore, connectivity modelling approaches are very diverse. To cover them all is beyond the scope of this introduction. We can mention that early efforts relied on metapopulation theory and island biogeography, and population genetics. This thesis focuses on population level modelling at the landscape scale. \\

\cite{fletcher_spatial_2018} emphasizes three common approaches to connectivity modelling: structural landscape modelling, resistance-based modelling and graph-based measures. Structural models are simple representations of physical landscape barriers. Resistance based approaches can go beyond structural approaches by abstracting the landscape into a resistance surface representing the level of difficulty with which animals move into the landscape. Its parameterization can depend on the life history and ecology of species, making it an appealing method for functional connectivity modeling.  It is the method used in this thesis: we derive resistance surfaces from an habitat suitability analysis. Graph based approaches depend on graph theory and are also common and often seen as complementary to resistance-based approaches. \\

Resistance surfaces can be used directly to calculate the shortest paths between patches with lowest cumulative resistance (the so-called “least-cost paths”), and derive species-specific networks based on thresholding \citep{lechner_characterising_2017}. An alternative method to treat resistance surfaces is to use circuit theory to estimate resistance distances by simulating random walks. This allows prediction of flow for entire landscapes. According to \cite{fletcher_spatial_2018}, this approach can prove to be useful because it accounts for the redundancy in paths \cite{mcrae_isolation_2006, mcrae_mcrae_2008}. It is also an appealing approach as it allows to identify “pinch-points” (bottlenecks of animal movement). \\

This thesis makes use of circuit theory to estimate change to potential functional connectivity in 5 umbrella species, under a certain number of land use change scenarios. Next, we introduce land use change modelling. \\ \vspace{2em}

@@

Depending on the scenario, the model generates different forest change dynamics. The Historic scenario assumes no change in forest composition and stands as our control. While the baseline scenario assumes no further warming and status quo forest change dynamics, the RCP 8.5 scenario is one of the most severe climate change scenarios, and the trends observed in the baseline scenario are exacerbated \citep{ipcc_summary_2013}. We also derive two simple land use change scenarios: Business as usual (BAU) and Reforestation. Those two scenarios are meant to represent two extremes in land use change, and it is likely that the most probable trajectories for our landscape will be within those two boundaries. In combination with the climate scenarios, this brings our total scenario number to 6. \\

What predictions can be made as to how connectivity will be affected under our model? There has been a significant research effort to study and predict forest dynamics under climate change in Quebec. It has been shown that the distributions of many of the most abundant trees occurring in the region are out of equilibrium with climate at the margin of their range \citep{talluto_extinction_2017}. Some papers points to existing and expected lags in the response of forest to climate change \citep{savage_elevational_2015}, and others have commented on how disturbances and forest management practices are likely to fasten and facilitate range shifts \citep{leithead_northward_2010, boulanger_climate_2019, vieira_paying_2020}. A major prediction that can be made is that although boreal type (coniferous) forest is likely to respond more slowly to climate range than deciduous type, both coniferous and mixed population types in southern Quebec are likely to become more rarer, and be replaced by deciduous stands. \\

@@

}%
\Intro

>>>>

\SetAcknowledgeText{
For providing vital guidance and parts of the code, I thank Dr. Bronwyn Rayfield. For providing help with computing resources parts of the code, I thank Dr. Guillaume Larocque. For providing advice on STSIM implementation and code, I thank Dr. Colin Daniel and Dr. Leonardo Frid. I thank Dr. Guillaume Larocque and Dr. Yann Boulanger for their help in providing the LANDIS outputs and editorial help for the associated appendix. \\
\indent I thank all students, staff and organizations involved in the design and realisation of the workshop. I thank Aurélie Sierra for her invaluable guidance and counsel. I thank all students and staff facilitators: Madeleine Gauthier, Mikayla Salmon-Beitel and Violette Bertrand. I thank Geneviève Poirier and Mariève Lafontaine for their participation in the design, planning and organization of the workshop. I also thank Renée Gagnon for her precious help securing a location for the workshop. \\
\indent For providing essential funding for this project, I thank the Fonds de recherche du Québec – Nature et technologies (FRQNT). For providing logistical support I thank Dr. Helen Elina and Mr. Philippe Auzel. \\
\indent For providing key information on land use data, I thank Martin Joly, Marie-Josée Côté and Benoit Jobin. \\
\indent I thank my supervisor, Prof. Andrew Gonzalez, for his support and for believing in my abilities every step of the way. I also thank members of my committee, Prof. Brian Leung and Prof. Elena Bennett for their timely and wise council at every committee meeting. I thank all my fellow grad students and especially Alex Arkilanian for all the help revising proposals, discussion ideas, and providing key support. \\
\indent I thank my mother Laura, my father Philippe, and Talia for their unwavering support. Thank you to Blanka and Peter for sheltering me and putting up with my writing process during the thick of the pandemic. \\
}%
\Acknowledge%

@@

% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetIntroName{{General Introduction}}
\SetIntroText{

Space is a finite resource: there is a limit on how much land is available for different human activities. How communities manage and govern space is a reflection of the choices and trade offs made by different people and organizations at different spatial and temporal scales, making the study of land systems a social, historical, and ecological discipline. The political and governance processes by which those trade-offs are resolved is of vital importance for biodiversity and ecosystem function. These processes ultimately determine and regulate land use and land cover change (LUC): if and how the resources held on the land are exploited, transformed or conserved. \\

LUC is a major cause of habitat loss and fragmentation, and one of the most important threats to biodiversity and ecosystem function worldwide \citep{newbold_global_2015}. It is estimated that LUC was responsible for more than one-third of animal extinctions between 1600 and 1992 \citep{centre_global_1992}. Moreover, a recent report from the IPBES (Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services) shows that about about 9\% of all described terrestrial species might suffer from habitat loss to the point that it threatens their long-term persistence \citep{ipbes_summary_2019}. Rarely, however, is habitat loss a complete process; once continuous tracts of habitat are often left as a mosaic of habitat fragments of different sizes separated by land cover with differing degrees of value to the resident fauna and flora.  Habitat fragmentation drives additional loss of diversity and ecosystem services because it reduces the connectivity of the landscape \citep{haddad_habitat_2015}. \\

Ecological connectivity is the extent to which the landscape supports the movements of organisms \citep{gonzalez_spatial_2017}, and is paramount for the resilience of both populations and ecosystem services \citep{mitchell_monteregie_2015} in heterogeneous and fragmented landscapes. Land use changes such as urban sprawl can cause deforestation, fragmenting habitats, and slowly eroding ecological connectivity. Many urban landscapes are experiencing uncontrolled urban sprawl and have suffered losses in connectivity and ecosystem services as a consequence. Examples include cities like Barcelona \citep{marulli_gis_2005}, New York City \citep{mcphearson_urban_2014}, and also Montréal \citep{dupras_urban_2015}.  The drivers behind those land use changes are complex and understanding them is an obstacle to conservation planning  \citep{worboys_connectivity_2010}, including connectivity conservation planning. \\

Connectivity conservation planning (CCP), refers to the enterprise that engages multiple actors such as academics, NGOs, governmental bodies at different scales and in a common goal to conserve and restore the ecological connectivity of the landscape. CCP methods usually involve modelling the connectivity of the landscape of interest and using a prioritization method to determine conservation priorities, and modify land use practices to conserve or improve connectivity. This thesis  contributes to an ongoing CCP in Montérégie, in southern Québec. In the following, we introduce connectivity modelling. We then introduce the study region and describe previous CCP and research. We point out ways in which CCP methods can be improved and how this thesis contributes to doing so. We then introduce the two other key concepts of the thesis: land use change modelling and participatory methods. \\ \vspace{2em}

Connectivity is a pillar of spatial ecology, the study of how ecological dynamics and interactions are structured in space. Connectivity is well known to influence various aspects of populations and community ecology, including dispersal, demographic rescue, and the spread of diseases. In conservation, connectivity is seen as a landscape feature that can be protected and restored to counteract the effects of  land use and climate change, and ensure the persistence of species in a landscape. \\

Connectivity as a concept has been used loosely in ecology and conservation. In their textbook “Spatial Ecology and Conservation Modeling”, \cite{fletcher_spatial_2018} identify up to 11 different variations on the concept of connectivity. For the focus of this thesis, it is important to make two important distinctions with regard to connectivity First, a distinction is often made between structural and functional connectivity. Structural connectivity emphasizes landscape structure and configuration, while functional connectivity attempts to capture species’s diverse responses to landscape’s configuration, and pays more attention to species-specific ways of moving across the landscape. Within functional connectivity another distinction is to be made between potential and realized connectivity \citep{calabrese_comparison_2004}. Potential connectivity estimates how much a landscape could facilitate or impede animal movements, based on indirect information, such as species-specific models and expert knowledge. Realized connectivity, on the other hand, relies on observational data to measure animal movement directly to quantify connectivity. This thesis focuses on modelling potential functional connectivity. \\

Connectivity modelling methods depend on the level of organization (individual, population, species or community) and the spatial scale (patch or  landscape) that connectivity is considered. To cover them all is beyond the scope of this introduction, but it is important to replace our choice of methodology in the larger body of foundational work on spatial ecological modelling. This thesis focuses on species level modelling at the landscape scale, and this focus has to be justified. Important efforts in the field of spatial ecological modelling relied on metapopulation theory, island biogeography (IBT), and population genetics. These bodies of theory spurred different approaches to modelling. For example, an IBT approach is focused on predicting diversity (number of species) in island patches \citep{macarthur_theory_1967} and has largely influenced other bodies of work in spat \citep{warren_how_2020}. On the other hand, metapopulation theory is focused on predicting long-term patch occupancy \citep{hanski_habitat_1999, hanski_metapopulation_2000}. Finally, population genetics focuses on predicting gene variation and flow within populations \citep{lewontin_population_1985, charlesworth_population_2017}. Unlike these methods, our goal is to estimate potential animal movement at the species level, over a large spatial and temporal extent. We therefore rely on a more appropriate approach which combines habitat suitability analysis and circuit theory modelling.\\

\cite{fletcher_spatial_2018} emphasizes three common approaches to connectivity modelling: structural landscape modelling, resistance-based modelling and graph-based measures. Structural models are simple representations of physical landscape barriers. Resistance based approaches can go beyond structural approaches by abstracting the landscape into a resistance surface representing the level of hindrance to animal movement into the landscape. The parameterization of resistance surfaces depends on the life history and ecology of a species, making it an appealing method for functional connectivity modeling.  This is the method used in this thesis: we derive resistance surfaces for multiple species as input to a habitat suitability analysis. Graph based approaches depend on graph theory and are also common and often seen as complementary to resistance-based approaches. \\

Resistance surfaces can be used directly to calculate the shortest paths between patches with the lowest cumulative resistance (the so-called “least-cost paths”), and derive species-specific networks based on resistance thresholding \citep{lechner_characterising_2017}. An alternative method to treat resistance surfaces is to use circuit theory to estimate resistance distances by simulating random walks. This allows prediction of flow for entire landscapes. According to \cite{fletcher_spatial_2018}, this approach can be useful because it accounts for the redundancy in paths \cite{mcrae_isolation_2006, mcrae_mcrae_2008}. It is also an appealing approach as it allows the identification of “pinch-points” (bottlenecks to animal movement). \\

This thesis makes use of circuit theory to estimate change to potential functional connectivity for 5 umbrella species, under a certain number of land use change scenarios. These 5 species were selected as their connectivity needs have been shown to be representative of the needs of a larger pool of 14 species (see \cite{albert_applying_2017, meurant_selecting_2018}). See table \ref{tab:species} for a detailed description of the 5 species. Next, we introduce the study region on which we carry these analyses. \\ \vspace{2em}

@@

Montérégie is one of the 17 administrative regions of the Canadian province of Québec. It is situated southeast of the city of Montréal, Québec, on the south bank of the Saint Lawrence, and is therefore located in the ecoregion of the Saint Lawrence Lowlands. Its surface area is \SI{11852}{\km\squared} and is home to about $1.5$ million inhabitants. Agriculture is the predominant land use in the region, and about half of this land area is covered by agricultural land \citep{isq_quebec_2020}. Statistique Québec indicates that between 1995 and 2008, 8\% of natural habitat and 4\% of agricultural lands has been artificialised. These changes are spread over the 15 MRC (“Municipalite Regionales de Comte”) making up Montérégie. MRCs are essentially sub-administrative regions at the level of which land use orientations are decided through two important documents: the PDZA (Plan de Développement de la Zone Agricole) and the SAD (Schéma d'Aménagement et de Développement). These MRCs are made up of 177 municipalities. In addition to MRCs, two First Nation reserves are present in Montérégie. \\

The eastern part of Montérégie is also under the jurisdiction of the Greater Montréal Area (GMA). The ecological connectivity of the GMA and its benefits as a provider of ecosystem services has recently been assessed in a report to the Québec ministry of the environment \citep{rayfield_priorisation_2018}. This study focused on identifying regions of highest ecological connectivity, and therefore of highest priority for the conservation of biodiversity and ecosystem services. Other work by Rayfield et al. (2019, unpublished) has extended the analysis of connectivity to the whole of the Saint Lawrence Lowlands. Prior to this research, the project of “Montérégie Connection”, spearheaded by an interdisciplinary team of researchers from McGill University, studied the provision of ecosystem services in the region and produced important knowledge for conservation planning. \\

Deriving insight from these contributions, multiple stakeholders have become mobilized to set connectivity conservation goals for the region. A number of Non-Governmental NGOs have designed corridor networks for their jurisdiction, with varying degrees of inputs from previous scientific work; these include Nature Action Québec (NAQ), Centre de la Nature du Mont Saint Hilaire (CDN), theCentre d'interprétation du milieu écologique du Haut-Richelieu (CIME), and Appalachian Corridors (CA). These associations are now beginning to work together to design science-driven connectivity conservation plans for the whole region, and are in need of science-based decision support tools. They are also concerned with the social acceptability of conservation, and with the resilience of their plans given expected future land use and climate change. \\ \vspace{2em}

@@

The maps produced by previous analyses represent a snapshot of the current state of connectivity in the region. But researchers are developing methods for including future land use and climate change impacts in their analyses. For example, \cite{albert_applying_2017}, used a simple land use and projection from regional climate change models to identify a network of protected habitat robust to climate change for a subset of Montérégie. This was an important step, because ignoring land use change could lead to ill-adapted conservation plans. However, connectivity conservation models rarely include both land use and climate change in their calculations \citep{costanza_landscape_2019}, which and so it remains a limitation in many science-based approaches to  CCP. \\

Similarly, current CCP methods often do not confront the results of the prioritization with the priorities perceived by stakeholders. There is a need for planning tools to allow for this interaction to happen. Those tools need to be able to guide stakeholders through an iterative process of conservation and land use scenario building. Methods are also available to include stakeholder inputs into CCP. Those “participatory research methods” are designed to conduct the research process “with those people whose life-world and meaningful actions are under study” \citep{bergold_participatory_2012}. Those methods often rely on collecting data through a community-driven process during workshops (for an example of a recent local participatory effort in Monteregie see \cite{martins_results_2020}). \\

In this thesis, we attempt to remedy those two limitations. In the first chapter, we combined a connectivity model using circuit theory with a land use and climate change model that uses a combination of statistical modelling and MCMC-based simulations. In the second chapter, we ran a stakeholder workshop to derive conservation priorities and use those results as inputs to derive simple, community driven, alternative conservation scenarios. We compare and contrast the results of these scenarios, and discuss the importance of considering stakeholder inputs to produce a resilient network of protected areas. \\

Next, we introduce the two other key conceptual themes of the thesis: land use change modelling and participatory methods. \\ \vspace{10em}

@@

Depending on the scenario, the model generates different forest change dynamics. The Historic scenario assumes no change in forest composition and stands as our control. While the baseline scenario assumes no further warming and status quo forest change dynamics, the RCP 8.5 scenario is one of the most severe climate change scenarios, and the trends observed in the baseline scenario are exacerbated \cite{IPCC}. We also derive two simple land use change scenarios: Business as usual (BAU) and Reforestation. Those two scenarios are meant to represent two extremes in land use change, and it is likely that the most probable trajectories for our landscape will be within those two boundaries. In combination with the climate scenarios, this brings our total scenario number to 6. \\

What predictions can be made as to how connectivity will be affected under our model? There has been a significant research effort to study and predict forest dynamics under climate change in Québec. It has been shown that the distributions of many of the most abundant trees occurring in the region are out of equilibrium with climate at the margin of their range \citep{talluto_extinction_2017}. Some papers points to existing and expected lags in the response of forest to climate change \citep{savage_elevational_2015}, and others have commented on how disturbances and forest management practices are likely to fasten and facilitate range shifts \citep{leithead_northward_2010, boulanger_climate_2019, vieira_paying_2020}. A major prediction that can be made is that although boreal type (coniferous) forest is likely to respond more slowly to climate range than deciduous forest, both coniferous and mixed population forests in southern Québec are likely to be replaced by deciduous stands. \\

@@

This thesis has two chapters.  In the first chapter, we combined a connectivity model, using circuit theory, with a land use and climate change model that uses a combination of statistical modelling and MCMC-based simulations. In the second chapter, we ran a stakeholder workshop to derive conservation priorities and use those results as inputs to derive simple, community driven, alternative conservation scenarios. We also compare and contrast the results of these scenarios and discuss the importance of considering stakeholder inputs to produce a resilient network of protected areas. \\
}%
\Intro
